Deploying and Sharing QSAR Models using Predictive Modeling Markup Language

### 2014-12-02 paper.tex
Broad outline of the "Discussion" section

>>>>

Putting the previous and current work of the ONS team in perspective.
The analysis could be based on the Melting Point dataset \url{http://onswebservices.wikispaces.com/meltingpoint}, which has practical 
applications in designing better laboratory experiments.

Evolution of the Melting Point dataset \url{http://onschallenge.wikispaces.com/Open+Melting+Point+Datasets}. Over time, has been analyzed as a complete set 
(over 10'000 chemical compounds) and chemical class-based subsets (few
hundred to few thousand compounds).

The preferred model training algorithm for the complete set has been
Random Forest. RF model is technically quite challenging to work with,
because it is typically several orders of magnite larger than other
common algorithms (eg. linear regression, SVM regression).

Notable accomplishments:
\begin{itemize}
  \item MPM001 \url{http://onschallenge.wikispaces.com/MeltingPointModel001}.
  Analysis and visualization of the chemical space using PCA, exploratory 
  modeling using RF vs. linear regression
  \item MPM002 \url{http://onschallenge.wikispaces.com/MeltingPointModel002}.
  Web service version of MPM001. PHP for frontend, R and CDK for backend.
  \marginnote{This looks like a rather tricky setup. Intended to be a single
  central installation that serves all interested parties?}
  \item MPM004 \url{http://onschallenge.wikispaces.com/MeltingPointModel004}.
  Using the high-quality subset (2706) of the complete set. The resulting RF
  model is available for download in R native RDS data format.
  \item MPM009 \url{http://onschallenge.wikispaces.com/MeltingPointModel009}
  QsarDB approach. The model is exported in RDS data format instead of PMML
  data format, but the resulting QDB archive contains CDK descriptor calculation
  information, which makes it executable by anyone using open source QsarDB 
  tools. Internally, the RF model data structure was converted to a generic
  wrapper object that exposed "public API" in the form of methods getPropertyId
  getDescriptorIds and evaluate(Map of descriptor valuess).
  \item MPM011 \url{http://onschallenge.wikispaces.com/MeltingPointModel011}.
  OpenBabel approach. Molecular descriptors are calculated using OpenBabel
  software as fragment counts. The model is exported in OpenBabel data format
  for fragment contribution models \url{http://onschallenge.wikispaces.com/file/view/mpC.txt/350881856/mpC.txt}. This model can be applied to any chemical structure
  representation that OpenBabel supports in input. This fragment contribution 
  table is the final model representation. The OpenBabel approach is applicable
  only to linear models (ie. counts of atoms/bonds/fragments multiplied by their
  contributions). Cannot do more complex model types, unless the developers
  of the OpenBabel project decide to work on it more.
  \item MPM012 \url{http://onschallenge.wikispaces.com/MeltingPointModel012}.
  QSAR.io approach (ONS workbook is not completed). The RF model is represented
  in PMML data format, and integrated with CDK descriptors over Java user-defined
  functions (UDFs). This will make it possible to evaluate such QSAR models
  in all Java environments, including real-time scoring, in-database scoring
  and batch scoring.
\end{itemize}

ONS Melting Point models assume 2D chemical structure representation as input
(eg. SMILES, InChI). It is possible to use 3D representations that are recognized
by CDK and can be converted to the same in-memory IChemObject form (eg. Sdf).
At later stages, ONS team began collaborating with ChemSpider and added support
for ChemSpider Identifiers (CSIDs). When given a CSID as input, then an ONS
webservice first resolves to an appropriate 2D representation by performing a
query against ChemSpider database. In the future, would be equally useful to 
support IUPAC systematic names.
